Direct observation of Feshbach enhanced \swave scattering of fermions

### 2015-06-19 draft1.tex
Went through most of the other authors' comments
Still have to do the pictures and another read-through

\begin{document}
\newcommand{\swave}[0]{$\it{s}$-wave }
\newcommand{\pwave}[0]{$\it{p}$-wave }
\newcommand{\K}{ $^{40}\rm{K}$}
\newcommand{\Rb}{ $^{87}\rm{Rb}$}
\newcommand{\us}{$\mu \rm{s}$} 
\newcommand{\mT}{$m\rm{T}$}

@@

\begin{abstract}
We directly measured the \swave{} scattering cross-section of ultracold \K{} atoms accross a Feshbach resonance. We collided a pair of degenerate clouds and imaged the scattered atoms. Owing to their low density, few atoms scattered, even near the resonance. To optimize signal to noise, we develop techniques to interpret absorption imaging in the regime where the optical intensity changes dramatically as light traverses the cloud, and recoil induced 
detuning corrections are significant. We applied these techniques to our \swave{} scattering data and extracted the resonant magnetic
field value. These imaging techniques are generally applicable, and can be used to observe effective \pwave{} scattering in the presence of spin-orbit coupling in a spin polarized  Fermi gas.

@@

Feshbach resonances are a stand-by technique for tuning the interaction strength in ultracold atomic gases. Even weak interactions play a crucial role in the physics of atomic Bose-Einstein Condensates (BECs), giving rise to, for example, their characteristic Thomas-Fermi density profiles. The effects of interactions in Fermi gases, however, are harder to observe. The density of Fermi clouds is typically reduced by a factor of $~10^3$ from that of BECs, making it necessary to enhance the strength of interactions in order to observe them. Feshbach resonances alter the inter-atomic scattering length via an externally applied uniform magnetic field. Not only do they enable detection of interactions, but they make it possible to tune interactions from attractive to repulsive, allowing for the phase transition from the BCS to BEC regime at sufficiently low temperatures \cite{RegalThesis, Regal04}. 
\par A Feshbach resonance occurs when a bound molecular state of few atoms energetically approaches the free atomic state \cite{Chin10, Timmermans99}. The relative energy of the free atomic states and the molecular state is defined by a bias magnetic field. The Feshbach resonance can thus be approached by changing the bias field. The effect of the resonance on the scattering length between two free atoms in the appropriate hyperfine states is

@@

\par To use Feshbach resonances as a tool it  is necessary to have a good measurement of the parameters in the above equation.  The exact value of the resonant field value $B_0$ is impossible to calculate analytically and must be estimated via numerical models \cite{Tiesinga93, Lysebo09, Gao11} or determined experimentally \cite{Inouye98, Cornish00}. Some experimental techniques that have been used to characterize Feshbach resonances include: the observation of loss due to three-body inelastic scattering, re-thermalization timescales, and anisotropic expansion, all of which infer the elastic scattering cross section from collective behavior of the cloud \cite{Regal03,OHara02,Monroe93}. 
\par Here we present a new technique for measuring the location and width of a Feshbach resonance. We collide a pair of ultra-cold Fermi gases and directly image the resulting \swave scattering halo as a function magnetic field strength. This allows us to observe the enhancement in scattering without relying on proxy effects. We measured the fraction of atoms scattered during the collision, and from this fraction deduced the resonant magnetic field  and width of the resonance.

@@

\par Our Fermi gases are so dilute that even with the resonant enhancement of the scattering cross section, only a small fraction of the atoms scatter. This makes direct detection of \swave scattering halos difficult due to detection uncertainty, which disproportionately affects regions of low atomic density. To optimize the signal to noise for low atom numbers, we utilized a non standard imaging regime. In this regime, simulation was necessary for an accurate interpretation of the images. We performed these simulations and used the results to extract the atom number and the scattered fraction from our images.
\par This paper is in two parts. In the first, we study absorption imaging in the presence of significant time-dependent Doppler shift and show how we use our results to interpret data. In the second, we describe our\swave scattering experiment and extract a measure of the location of the Feshbach resonance in \K{}.

@@

Every experiment requires a detection scheme. In ultracold atomic experiments, most measurements are based on interacting laser light with the atomic cloud. In our experments we used absorption imaging, the most commonly used technique. Absorption imaging relies on optical transitions between the ground and some selected excited atomic states. Such atomic transitions have an energy difference $E_0$ with an associated frequency $\omega_0 = E_0/\hbar$, and a natural transition linewidth $\Gamma$. When interacting with a laser light field an atom will scatter photons: it will absorb a photon from the light field and move up to its excited state,  then re-emit the photon and decay back to its ground state. The  rate at which this scattering occurs is

@@

\par An absorption image is obtained by shining an on or near resonant probe beam ($\delta\approx0$) onto the atomic cloud. Some of the light is scattered by the atoms, and the part of the beam that made it through the cloud, $I_f(x,y)$ is imaged onto a camera, as seen in Fig. \ref{fig:absorptionIntor}a (top). After the atoms leave the imaging volume, the probe light is reapplied to calibrate the intensity $I_0(x,y)$ of light unaffected by the atoms (bottom). 

@@

\caption{Absorption imaging. a. Near resonant probe light illuminates the atoms, and the transmitted light (containing a shadow of the atoms) is imaged on the camera. After the atoms depart an image of just the probe light is taken. b.  The probe beam is partially absorbed as it traverses the cloud, and the intensity seen by atoms further along the imaging direction $z$ is lowered.  c. An atomic cloud illuminated by a probe light field absorbs photons from the probe and re-emits them in all directions. This process results in a net velocity of the cloud in the direction of the probe light as well as diffusive spreading in the transverse directions.  }  

@@

\par These intensities are related to the number of atoms the light encountered. Consider the light as it travels along the imaging axis $e_z$ through a 3-D atomic density profile $\rho(x,y,z)$. For the purposes of this discussion, let us focus on a single pixel of the camera, and thus a single value of $I_0$ and $I_f$, and a single column of atoms, $\rho(z)$. We will not be able to infer the entire atomic distribution, but for each pixel we can obtains the integrated density, $n = \int \rho\left(z\right) \,\mathrm{d}z$. As the light travels through a column of atoms, each atom will scatter light according to Eq. (\ref{eq:scatrate}). Therefore, the atoms further along the imaging axis $z$ will experience a reduced optical intensity, as seen in Fig. \ref{fig:absorptionIntor}b. On resonance ($\delta=0$), the intensity change from scattering as a function of $z$ is

@@

\frac{dI(z)}{dz}=\hbar\omega_L\rho\gamma_{sc}=-\rho\sigma_0\frac{I(z)}{1+I(z)/I_{\rm{sat}}},

@@

 This equation can be easily integrated to obtain  \cite{Reinaudi07}

@@

In the limit where the probe intensity is much smaller than the saturation intensity, $I_0\ll I_{\rm{sat}}$, this reduces to the optical depth, $OD_0=-\ln \left(I_f/I_0\right)$. Since this is the simplest possible model that relates the observed intensities to the column density $n$. When the probe intensity is comparable to the saturation intensity, the second term in Eq. (\ref{eq2}) becomes significant, and we define the optical density corrected for high probe intensity
OD_1 = OD_0 + \frac{I_0-I_f}{I_{\rm{sat}}}.
\label{eq:OD1}
There are further corrections that this equation does not take into account. In particular, it neglects the atomic recoil momentum and its effect on the laser detuning \cite{Konstantinidis12}, which will be discussed below.
\par When an atom absorbs a photon from the laser light field and is excited to a higher energy level, by conservation of momentum it must also acquire a velocity in the direction of the light field. This recoil velocity is given by $v_r=\hbar k/m$, where $k=2\pi/\lambda$ is the wavenumber of the light field with wavelength $\lambda$ and $m$ is the atomic mass. When the atom returns to its ground state, the photon is re-emitted with some momentum $\vec{p}_e$. Over many photons, this momentum distribution averages to zero, $\int\vec{p}_e\mathrm{d}^3 x=0$.  The variance, however, is not zero, allowing the atoms to acquire some momentum transverse to the laser field. We will ignore this correction, but the effect of this on the atomic cloud is pictured in Fig. \ref{fig:absorptionIntor}c. 

@@

 However, as shown in  Fig. \ref{fig:ODcorrections}, the perturbative treatment breaks down shortly after the zeroth order approximation of Eq. (\ref{eq2}). To adequately correct for the recoil induced detuning of the atoms, we numerically simulated the imaging process for $I_f$ as a function of imaging time, atomic density, and probe intensity. 

@@

\par In the following, we describe two versions of this simulation. First, we took a simplistic approach where the spatial distribution of atoms does not change appreciably during the imaging time: $vt\ll I_0/(\hbar \omega_L \gamma_{sc} \rho)$. We tested this approach in known limits and then checked the validity of the static assumption. For realistic input parameters, we found this assumption to be invalid. We developed a slightly more sophisticated quasi-classical approach and allowed the atoms to move during the imaging time, allowing us to simulate the phase space evolution of atoms subjected to probe light. While the atomic trajectories were wildly different than in the static atom approximation, the predicted $OD$s only varied on the 0.5$\%$ level between the two models.

\subsection{Stationary atom model}
To solve Eqs. (\ref{eq3})-(\ref{eq4}), we started with a 1-d distribution of atomic densities $\rho(z)$, assumed to be Gaussian in shape. We divided the cloud into spatial bins (the bin size was chosen such that decreasing the bin size further changed the result at the $10^{-5}$ level).  In this approximation, we kept the number of atoms in each bin constant.  The algorithm used is shown in Alg. [\ref{algorithm1}]. We call the optical depth obtained from this algorithm $OD_{corr1}$.

@@

\par We checked the validity of our simulation in the limits where the problem is analytically solvable. When the probe intensity is much weaker than the saturation intensity, $I_0\ll I_{\rm{sat}}$, the atoms' velocity is hardly changed, and Eq.(\ref{eq3}) reduces to 
\begin{equation}
\frac{dI(z)}{dz}=-\rho\sigma_0 I(z),
\end{equation}
from which we recover 
\begin{equation}
\sigma_0 n = OD_0 = -\ln I_0/I_f. \label{eq6}
\end{equation}
In the limit that the probe intensity is much stronger than the saturation intensity, $I_0/I_{\rm{sat}}\gg \delta/2\Gamma$, even far detuned atoms will absorb light at their maximum, allowing us to again neglect the time dependence of the detuning. Equation (\ref{eq3}) becomes 
\begin{equation}
\frac{dI(z)}{dz}=-\rho\sigma_0 I_{\rm{sat}}, 
\end{equation}
which integrates to 
\begin{equation}
\sigma_0 n = \frac{I_0 - I_f}{I_{\rm{sat}}}. \label{eq8}
\end{equation}
We recognize the right hand sides of Eq. (\ref{eq6}) and Eq. (\ref{eq8}) as the two terms in the expression for $OD_1$ in Eq. (\ref{eq2}). Thus, as shown in  Fig. \ref{fig:IsatLimits}, in both limits $OD_{corr1}$  coincides with $OD_1$ in both the small and large probe intensity limits. 
\begin{figure}
	\includegraphics{figure3.pdf}
\caption{Optical depth as a function of probe intensity as predicted by the simulation (blue symbols) and by Eq. (\ref{eq6}) (green curves), for three different imaging times. The predictions agree in both the high and low intensity limits, and differ for probe intensities comparable to the saturation intensity and longer imaging times. }  
\label{fig:IsatLimits}
\end{figure}
\par We used the results of this simulation to check if the stationary atom assumption is valid, i.e. if the distance traveled by the atoms during the imaging time is less than the bin size. As can be seen from Fig. \ref{fig:atomTravel}, not only do the atoms travel more than the bin size, but they travel far beyond the initial extent of the cloud. Moreover, owing to the high scatter rate, the back of the cloud overtakes the front for long imaging times. Thus, the atomic distribution as a function of position changes dramatically during the imaging pulse, and the stationary assumption is invalid. 
\begin{figure}
	\includegraphics{figure4.pdf}
\caption{Position of atoms as a function of imaging time for atoms in the first, middle, and last bins of the simulation, obtained by integrating their velocities. The probe intensity here is $1.2 I_{\rm{sat}}$, and the optical depth is 1.6.}  
\label{fig:atomTravel}
\end{figure}

@@

To account for the changing atomic distribution during the imaging pulse, we simulated the classical kinetics of atoms subject to the recoil driven optical forces. To simulate large ensembles in a reasonable time, we used ``superatoms",  each describing the aggregate behavior of $N_{sa}$ atoms. The amended algorithm is shown in Alg.[\ref{algorithm2}]. 

@@

\STATE $z[n]=z_0$, $\delta[n]=0$ \COMMENT{initialize position and detuning for each superatom, labeled by index $n$}
\STATE $O[i]=n$ \COMMENT{make a list of superatom indexes, ordered by position}

@@

\STATE $n=O[i]$ \COMMENT{apply probe intensity to superatoms in order of appearance}
 \STATE $A=\sigma_0 N_{sa} dz$ \COMMENT{dz is length over which atoms were grouped into single superatom}

@@

\STATE $O[i]$=sort($n$, key =$z[n]$) \COMMENT{sort superatoms indexes by current position}

@@

\par First, we checked the velocity predicted in this model against known limits. One such limit is that of a single superatom. In this case, there is no attenuation, and the intensity seen by the superatom is constant at $I_0$. Only the detuning of the single superatom evolves in time, and Eqs. (\ref{eq3}) and (\ref{eq4}) give

@@

\caption{The velocity of a single superatom as a function of probe intensity for various imaging times. Simulation data (dots) and analytical solutions (lines) are in good agreement.}  

@@

\caption{Phase space evolution of an atomic cloud exposed to probe light with intensity $I_0=1.2 I_{\rm{sat}}$. We defined $\Delta v=v -\left< v(t) \right>$  and $\Delta z=z-\left< z(t) \right>$, subtracting out the center of mass position and velocity of the cloud. The optical depth is 1.6, and the initial cloud is a gaussian with a width of 10 $\mu$m in (a) and 1 $\mu$m in (b).}  

@@

\par We compared the optical depths predicted by each of the two models, $OD_{corr1}$ and $OD_{corr2}$. As seen Fig. \ref{fig:atomTravel}, the predicted optical depths were hardly changed by including the full time evolution:  $\left|OD_{corr1}-OD_{corr2}\right|/OD_{corr1} \le 0.005$. Thus, for the purposes of deducing the atom density from experimental optical depths,the stationary atom model is sufficient. Furthermore, we simulated a range of initial density profiles $\rho(z)$, and found their impact to be negligible - the only observable is the integrated atomic density $n=\int\rho(z)\mathrm{d}z$. 

@@

\label{fig:atomTravel}

@@

\subsection{Signal to noise optimization}
This simulation allowed us to interpret experimentally obtained initial and final intensities. For a given imaging time, we created a look-up table of predicted optical depth as a function of probe intensity and atomic density. We then found the observed optical depth on this table, with the given probe intensity, and inferred the atomic density. The uncertainty in the measured intensities can be propagated through this procedure, and we established optimal imaging parameters to maximize the signal to noise ratio of this detection scheme. 
\par Here, the only source of measurement uncertainty we considered was the Poisson distributed photon shot noise, with a standard deviation proportional to $\sqrt{N_p}$, where $N_p$ is the photon number. We then propagated this uncertainty through our correction scheme to obtain the uncertainty in our deduced value of $\sigma_0 n$. We define the signal to noise ratio (SNR) as $\sigma_0 n/\delta_{\sigma_0 n}$, where $ \delta_{\sigma_0 n}$ is the propagated measurement uncertainty.
\par As seen in Fig. \ref{fig:SNR}a, after about 40\us{} extending the imaging time no longer yields appreciable improvement in signal to noise ratio. Imaging for 40\us{} as opposed to 10\us{} where the uncorrected model is appropriate, improves the SNR by a factor of  1.5. We performed the experiments described in the second section at 40\us{} imaging time. Figure \ref{fig:SNR}b shows that the optimal probe intensity varies with the magnitude of $OD$s. For low atom numbers, $OD\approx0.1$, a probe intensity of $I_0\approx0.6I_{\rm{sat}}$ is best. However, in our experiment the probe intensity had a gaussian profile and was not uniform over the whole image.  The typical probe intensities used in our experiments were in the $I_0=0.1I_{\rm{sat}}-0.7I_{\rm{sat}}$  range.
\begin{figure}
	\subfigure[]{\includegraphics{figure9a.pdf}}
	\subfigure[]{\includegraphics{figure9b.pdf}}
\caption{Signal to noise ratio (SNR) for three different optical depths after correcting for recoil induced detuning. (a) SNR as a function of probe intensity $I_0/I_{\rm{sat}}=5.0$ and (b) SNR as a function of probe intensity for an imaging time of 50 \us{}.}  
\label{fig:SNR}
\end{figure}

>>>>

\begin{document}
\newcommand{\swave}[0]{$\it{s}$-wave }
\newcommand{\pwave}[0]{$\it{p}$-wave }
\newcommand{\K}{ $^{40}\rm{K}$}
\newcommand{\Rb}{ $^{87}\rm{Rb}$}
\newcommand{\us}{$\mu \rm{s}$} 
\newcommand{\mT}{$m\rm{T}$}
\newcommand{\ez}{$\hat{e}_z$}
\newcommand{\um}{$\mu \rm{m}$}

@@

\begin{abstract}
We directly measured the relative \swave{} scattering cross-section of ultracold \K{} atoms accross the 20.2 \mT{} Feshbach resonance. We collided a pair of degenerate Fermi clouds and imaged the scattered atoms. Owing to their low density, few atoms scattered, even near the resonance. To optimize signal to noise, we developed techniques to interpret absorption imaging in the regime where the optical intensity changes dramatically as light traverses the cloud, and recoil induced 
detuning corrections are significant. By applying these techniques to our \swave{} scattering data we were able to extract the resonant magnetic
field value. These imaging techniques are generally applicable to experimets requiring accurate detection of low atomic densities.

@@

Feshbach resonances are a commonly used technique for tuning the interaction strength in ultracold atomic gases. Even weak interactions play a crucial role in the physics of atomic Bose-Einstein Condensates (BECs), for example giving rise to their characteristic Thomas-Fermi density profiles \cite{KetterleBEC}. The effects of interactions in Fermi gases, however, are harder to observe. The density of Fermi clouds is typically $\sim1000$ times less than that of BECs, making it necessary to enhance the strength of interactions in order to observe them \cite{KetterleDFG}. Feshbach resonances alter the inter-atomic scattering length via an externally applied uniform magnetic field. Tuning this field to either side of the resonance changes the interactions from attractive to repulsive, allowing for the phase transition from the BCS to BEC regime at sufficiently low temperatures \cite{RegalThesis, Regal04}. 
\par A Feshbach resonance occurs when a few-atom bound molecular state energetically approaches the free atomic state \cite{Chin10, Timmermans99}. The relative energy of the free atomic states and the molecular state is defined by a bias magnetic field. The Feshbach resonance can thus be approached by changing the bias field. The effect of the resonance on the scattering length between two free atoms is \cite{Chin10}

@@

\par  The exact value of the resonant field $B_0$ is impossible to calculate analytically and must be estimated via numerical models \cite{Tiesinga93, Lysebo09, Gao11} or determined experimentally \cite{Inouye98, Cornish00}. Many experimental techniques have been used to characterize Feshbach resonances, including the observation of atom loss due to three-body inelastic scattering, measurement of re-thermalization timescales, and imagin anisotropic expansion of the cloud upon release from a confining potential, all of which infer the elastic scattering cross section from collective behavior of the cloud \cite{Regal03,OHara02,Monroe93}. 
\par Here we present a new technique for measuring the location and width of a Feshbach resonance. We collided a pair of ultra-cold Fermi gas clouds and directly imaged the resulting \swave scattering halo as a function magnetic field strength. This allowed us to observe the enhancement in scattering without relying on proxy effects. We measured the fraction of atoms scattered during the collision, and from this fraction deduced the resonant magnetic field  and width of the resonance.

@@

\par In our dilute Fermi clouds, even with the resonant enhancement of the scattering cross section, only a small fraction of the atoms underwent a scattering event as the clouds passed through each other. This made direct detection of \swave scattering halos difficult due to detection uncertainty, which disproportionately affected regions of low atomic density in absorption images. To optimize the signal to noise for low atom numbers, we utilized a non standard imaging regime. In this regime, the atoms acquire a non-negligible velocity during imaging and are Doppler-shifted away from the light field used to detect them.  Simulation of the absorption imaging process was necessary for an accurate interpretation of these images. With the simulation corrected images, we were able to extract the number and the scattered fraction of the atoms.
\par This paper is in two parts. In the first, we study absorption imaging in the presence of significant time-dependent Doppler shift and show how we use our results to interpret data. In the second, we describe our\swave scattering experiment and extract a measure of the location and width of the Feshbach resonance in \K{}.

@@

Absorption imaging is commonly thought of as the atomic cloud casting a "shadow" on a camera when illuminated by laser light. This imaging scheme relies on optical transitions between the ground and certain excited atomic states. Such atomic transitions have an energy difference $E_0$ with an associated frequency $\omega_0 = E_0/\hbar$, and a natural transition linewidth $\Gamma$. When interacting with a laser light field an atom will scatter photons: it will absorb a photon from the light field and move up to its excited state,  then re-emit the photon and decay back to its ground state. The  rate at which this scattering occurs, in the two-level atom approximation, is \cite{LCT}

@@

\par An absorption image is obtained by shining an on or near resonant probe beam ($\delta\approx0$) onto the atomic cloud. Some of the light is scattered by the atoms, and the unscattered light, $I_f(x,y)$, is imaged onto a camera, as seen in Fig. \ref{fig:absorptionIntor}a (top). After the atoms leave the imaging volume, the probe light is reapplied to calibrate the intensity $I_0(x,y)$ of light unaffected by the atoms (bottom). 

@@

\caption{Absorption imaging. a. Near resonant probe light illuminates the atoms, and the transmitted light (containing a shadow of the atoms) is imaged on the camera. After the atoms depart an image of just the probe light is taken. b.  The probe beam is partially absorbed as it traverses the cloud, and the intensity seen by atoms further along the imaging direction \ez{} is lowered.  c. An atomic cloud illuminated by a probe light field absorbs photons from the probe and re-emits them in all directions. This process results in a net velocity of the cloud in the direction of the probe light as well as diffusive spreading in the transverse directions.  }  

@@

\par The intensity $I_f(x,y)$ is related to the number of atoms the light encountered. Consider the light as it travels along the imaging axis $e_z$ through a 3-D atomic density profile $\rho(x,y,z)$. For the purposes of this discussion, let us focus on a single pixel of the camera, and thus a single value of $I_0$ and $I_f$, and a single column of atoms, $\rho(z)$. We are not able to infer the entire atomic distribution, but for each pixel we can obtain the integrated density, $n = \int \rho\left(z\right) \,\mathrm{d}z$. As the light travels through a column of atoms, each atom scatters light according to Eq. (\ref{eq:scatrate}). Therefore, the atoms further along the imaging axis $z$ experience a reduced optical intensity due to attenuation by the other atoms, as seen in Fig. \ref{fig:absorptionIntor}b. On resonance ($\delta=0$), the intensity change from scattering as a function of $z$ is

@@

\frac{dI(z)}{dz}=\hbar\omega_L\rho(z)\gamma_{sc}(z)=-\rho(z)\sigma_0\frac{I(z)}{1+I(z)/I_{\rm{sat}}},

@@

 This equation is integrated to obtain  \cite{Reinaudi07}

@@

In the limit where the probe intensity is much smaller than the saturation intensity, $I_0\ll I_{\rm{sat}}$, this reduces to the optical depth, $OD_0=-\ln \left(I_f/I_0\right)$   \cite{Reinaudi07}. This is the simplest possible model that relates the observed intensities to the column density $n$. When the probe intensity is comparable to the saturation intensity, the second term in Eq. (\ref{eq2}) becomes significant, and we define the optical density corrected for high probe intensity $OD_1$ as the right hand side of Eq.  (\ref{eq2}).
\par There are further corrections that this equation does not take into account. In particular, it neglects the atomic recoil momentum and its effect on the laser detuning \cite{Konstantinidis12}. When an atom absorbs a photon from the laser light field and is excited to a higher energy level, by conservation of momentum it must also acquire a velocity increment in the direction of the light field. This recoil velocity is given by $v_r=\hbar k/m$, where $k=2\pi/\lambda$ is the wavenumber of the light field with wavelength $\lambda$ and $m$ is the atomic mass. When the atom returns to its ground state, the photon is re-emitted with some momentum $\vec{p}_e$. Over many photons, this momentum distribution averages to zero, $\int\vec{p}_e\mathrm{d}^3 x=0$. Therefore, on average, the atom  will acquire a net velocity of $v_r$ in the \ez{} direction.  The variance of $\vec{p}_e$, however, is not zero, allowing the atoms to acquire some momentum transverse to the laser field. We will ignore this correction, but the effect of this on the atomic cloud is pictured in Fig. \ref{fig:absorptionIntor}c. 

@@

 However, as shown in  Fig. \ref{fig:ODcorrections}, the perturbative treatment breaks down shortly after the zeroth order approximation of Eq. (\ref{eq2}). To adequately correct for the recoil induced detuning of the atoms, we numerically simulated the imaging process to obtain $I_f$ as a function of imaging time, atomic density, and probe intensity. 

@@

\par In the following, we describe two versions of this simulation. First, we took a simplistic approach where the spatial distribution of atoms does not change appreciably during the imaging time: $vt\ll I_0/(\hbar \omega_L \gamma_{sc} \rho)$. We tested this approach in known limits and then checked the validity of the static assumption. For realistic input parameters, we found this assumption to be invalid. We developed aquasi-classical approach and allowed the atoms to move during the imaging time, allowing us to simulate the phase space evolution of atoms subjected to probe light. While the atomic trajectories were wildly different than in the static atom approximation, the predicted $OD$s only varied on the 0.5$\%$ level between the two models.

\subsection{Stationary atom model}
To solve Eqs. (\ref{eq3})-(\ref{eq4}), we started with a 1-D distribution of atomic densities $\rho(z)$, assumed to be Gaussian in shape. We divided the cloud into spatial bins (the bin size was decreased until the result only differed at the $10^{-5}$ level from smaller bin sizes).  In this approximation, we kept the number of atoms in each bin constant.  The algorithm used is shown in Alg. [\ref{algorithm1}]. We call the optical depth obtained from this algorithm the `corrected' optical depth, $OD_{corr1}$.

@@

\par We checked the validity of our simulation in the limits where the problem is analytically solvable. In the limit where the probe intensity is much weaker than the saturation intensity, $I_0\ll I_{\rm{sat}}$, the atoms' velocity is hardly changed, and Eq.(\ref{eq3}) reduces to 
\begin{equation}
\frac{dI(z)}{dz}=-\rho\sigma_0 I(z),
\end{equation}
from which we recover the analytic form 
\begin{equation}
\sigma_0 n = OD_0 = -\ln I_0/I_f. \label{eq6}
\end{equation}
In the limit that the probe intensity is much stronger than the saturation intensity, $I_0/I_{\rm{sat}}\gg \delta/2\Gamma$, even far detuned atoms will absorb light at their maximum. The time dependence of the detuning can thus be neglected, and Eq. (\ref{eq3}) becomes 
\begin{equation}
\frac{dI(z)}{dz}=-\rho\sigma_0 I_{\rm{sat}}, 
\end{equation}
which integrates to 
\begin{equation}
\sigma_0 n = \frac{I_0 - I_f}{I_{\rm{sat}}}. \label{eq8}
\end{equation}
We recognize the right hand sides of Eq. (\ref{eq6}) and Eq. (\ref{eq8}) as the two terms in the expression for $OD_1$ in Eq. (\ref{eq2}). Thus, as shown in  Fig. \ref{fig:IsatLimits}, in both limits $OD_{corr1}$  coincides with $OD_1$ in both the small and large probe intensity limits. 
\begin{figure}
	\includegraphics{figure3.pdf}
\caption{Optical depth as a function of probe intensity as predicted by the simulation (blue symbols) and by Eq. (\ref{eq2}) (green curves), for three different imaging times. The predictions agree in both the high and low intensity limits, and differ for probe intensities comparable to the saturation intensity and longer imaging times. }  
\label{fig:IsatLimits}
\end{figure}
\par We used the results of this simulation to check if the stationary atom assumption is valid, i.e. if the distance traveled by the atoms as deduced from integrating the acquired recoil velocity over the imaging time is less than the bin size. As can be seen from Fig. \ref{fig:atomTravel}, not only do the atoms travel more than the bin size, but they travel far beyond the initial extent of the cloud. Moreover, owing to the high scatter rate, the back of the cloud overtakes the front for long imaging times. Thus, the atomic distribution as a function of position changes dramatically during the imaging pulse, and the stationary assumption is invalid. 
\begin{figure}
	\includegraphics{figure4.pdf}
\caption{Position of atoms as a function of imaging time for atoms in the first, middle, and last bins of the simulation for an initial cloud 50 \um{} in extent, obtained by integrating their velocities. The probe intensity used in this calculation is $1.2\, I_{\rm{sat}}$, and the optical depth is 1.6.}  
\label{fig:atomTravel}
\end{figure}

@@

To account for the changing atomic distribution during the imaging pulse, we numerically simulated the classical kinetics of atoms subject to the recoil driven optical forces. To simulate large ensembles in a reasonable time, composite atoms described the aggregate behavior of $N_{ca}$ atoms. The amended algorithm is shown in Alg.[\ref{algorithm2}]. 

@@

\STATE $z[n]=z_0$, $\delta[n]=0$ \COMMENT{initialize position and detuning for each composite atom, labeled by index $n$}
\STATE $O[i]=n$ \COMMENT{make a list of composite atom indexes, ordered by position}

@@

\STATE $n=O[i]$ \COMMENT{apply probe intensity to composite atoms in order of appearance}
 \STATE $A=\sigma_0 N_{sa} dz$ \COMMENT{dz is length over which atoms were grouped into single composite atom}

@@

\STATE $O[i]$=sort($n$, key =$z[n]$) \COMMENT{sort composite atom indexes by current position}

@@

\par To validate our code, we checked the velocity predicted in this model against known limits. One such limit is that of a singlecomposite atom. In this case, there is no attenuation, and the intensity seen by the composite atom is constant at $I_0$. Only the detuning  evolves in time, and Eqs. (\ref{eq3}) and (\ref{eq4}) give

@@

\caption{The velocity of a single superatom as a function of probe intensity for various imaging times. Simulation data (dots) and numerical  solutions of Eq. (\ref{eq10}) (lines) are in good agreement.}  

@@

\caption{Phase space evolution of an atomic cloud exposed to probe light with intensity $I_0=1.2\, I_{\rm{sat}}$. We defined $\Delta v=v -\left< v(t) \right>$  and $\Delta z=z-\left< z(t) \right>$, subtracting out the center of mass position and velocity of the cloud. The optical depth is 1.6, and the initial cloud is a gaussian with a width of 10 $\mu$m in (a) and 1 $\mu$m in (b).}  

@@

\par We compared the optical depths predicted by each of the two models, $OD_{corr1}$ and $OD_{corr2}$. As seen Fig. \ref{fig:compareModels}, the predicted optical depths were hardly changed by including the full time evolution:  $\left|OD_{corr1}-OD_{corr2}\right|/OD_{corr1} \le 0.005$. Thus, for the purposes of deducing the atom density from experimental optical depths, the stationary atom model is sufficient. Furthermore, we simulated a range of initial density profiles $\rho(z)$, and found their impact to be negligible \--- the only observable is the integrated atomic density $n=\int\rho(z)\mathrm{d}z$. For interpreting experimental images, we used the data generated by the travelling atom simulation.

@@

\label{fig:compareModels}

@@

\subsection{Signal to noise optimization}
This simulation allowed us to interpret experimentally obtained final intensities. For a given imaging time, we created a look-up table of predicted optical depth as a function of probe intensity and atomic density. We then found the observed optical depth on this table, with the given probe intensity, and inferred the atomic density. The uncertainty in the measured intensities can be propagated through this procedure, and we established optimal imaging parameters to maximize the signal to noise ratio of this detection scheme. 
\par Here, the only source of measurement uncertainty we considered was the Poisson distributed photon shot noise, with a standard deviation proportional to $\sqrt{N_p}$, where $N_p$ is the photon number. We then propagated this uncertainty through our correction scheme to obtain the uncertainty in our deduced value of $\sigma_0 n$. We define the signal to noise ratio (SNR) as $\sigma_0 n/\delta_{\sigma_0 n}$, where $ \delta_{\sigma_0 n}$ is the propagated measurement uncertainty.
\par As seen in Fig. \ref{fig:SNR}a, after about 40 \us{} extending the imaging time no longer yields appreciable improvement in SNR. Imaging for 40 \us{} as opposed to 10 \us{} where the uncorrected model is appropriate, improves the SNR by a factor of  1.5. We performed the experiments described in the second section at 40 \us{} imaging time. Figure \ref{fig:SNR}b shows that the optimal probe intensity varies with the magnitude of $OD$s. For low atom numbers, $OD\approx0.1$, a probe intensity of $I_0\approx0.6\,I_{\rm{sat}}$ is best. However, in our experiment the probe intensity had a gaussian profile and was not uniform over the whole image.  The typical probe intensities used in our experiments varied over the $I_0=0.1\,I_{\rm{sat}}-0.7\,I_{\rm{sat}}$  range.
\begin{figure}
	\subfigure[]{\includegraphics{figure9a.pdf}}
	\subfigure[]{\includegraphics{figure9b.pdf}}
\caption{Signal to noise ratio (SNR) for three different optical depths after correcting for recoil induced detuning. (a) SNR as a function of probe intensity $I_0/I_{\rm{sat}}=5.0$ and (b) SNR as a function of probe intensity for an imaging time of 50 \us{}.}  
\label{fig:SNR}
\end{figure}

\subsection{Calibration of saturation intensity}
A charge-coupled device (CCD) camera was used to take the absorption images.  Each camera pixel  converts the photons it is exposed to, with some efficiency, into photoelectrons, and digitally outputs an integer, called `counts', that is proportional to the photoelectrons.  However, the proportionality constant depends on many factors, such as the quantum efficiency of the camera, the electronic gain during the readout process, and the polarization of the probe light. 
\par We determined this proportionality constant through direct measurement. In the limit where the system is adequately described by the uncorrected $OD_0$, only the ratio of the initial and final intensities matter, and this proportionality constant is irrelevant. In all other regimes, however, the ratio of the initial and final intensities to the saturation intensity also comes into play, making the proportionality constant significant. One way to approach this calibration is to determine the value of the saturation intensity in `counts' per unit time. 
\par To calibrate the saturation intensity in camera counts per unit time, we took absorption images of a cloud of \K{} atoms at three different imaging times, 40 \us{}, 100 \us{}, and 200 \us{}, at varying probe intensities. In a small region at the center of the cloud the atomic density is approximately uniform, and we averaged the initial and final intensities of each pixel in that region. Thus, for each image we obtained $I_0$ and $I_f$, in counts per microsecond. We then did a least squares fit of $OD_{corr}$, our simulated optical depth, to the data at all three imaging times simultaneously. The two fit parameters were the atomic density $n$ at the center of the cloud and the value of $I_{\rm{sat}}$ in counts per microsecond. As seen in Fig. \ref{fig:isatCalib}, the model produced a good fit to the experimental data, and we obtained a calibration of the saturation intensity for our experiment. 
\begin{figure}
	\includegraphics{figure8.pdf}
\caption{The optical depth as a function of probe intensity for three imaging times: $t=40$\us{} (blue),  $t=75$\us{} (green),  $t=100$\us{} (red). The dots represent experimental data and the lines represent the best fit of simulated data. The optimal fit parameters pictured are a $\sigma_0 n$ of 1.62 and saturation intensity of 29 counts/\us{}. }  
\label{fig:isatCalib}
\end{figure}
